---
title: "Getting Started with PrakKomstatPert8"
author: "Finesya Aura Nadifa"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with PrakKomstatPert8}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

PrakKomstatPert8 is an R package developed to evaluate the assumptions of linear regression models. Before interpreting the coefficients of a regression model, it is important to verify that the underlying assumptions are satisfied.

The package provides several diagnostic functions that perform:

- Normality test of residuals
- Multicollinearity assessment
- Heteroscedasticity test
- Detection of influential observations
- Overall diagnostic summary

This vignette demonstrates the complete workflow for using PrakKomstatPert8.

# Loading the Package

```{r}
library(PrakKomstatPert8)
```

# Example Dataset

This vignette uses the built-in `mtcars` dataset available in R.

```{r}
head(mtcars)
```

The response variable is:

- `mpg` (Miles per Gallon)

Predictor variables:

- `wt` (Weight)
- `hp` (Horsepower)
- `disp` (Displacement)

# Fitting a Linear Regression Model

A linear regression model is fitted before performing diagnostic checks.

```{r}
fit <- lm(mpg ~ wt + hp + disp, data = mtcars)
summary(fit)
```

# Residual Normality

Residual normality is evaluated using the Shapiro–Wilk test.

```{r}
normality_check(fit)
```

### Interpretation

- **p-value > 0.05** indicates residuals are approximately normally distributed.
- **p-value ≤ 0.05** indicates evidence against normality.

# Multicollinearity

Multicollinearity occurs when predictor variables are highly correlated. 
PrakKomstatPert8 computes the Variance Inflation Factor (VIF) manually.

```{r}
multicol_check(fit)
```

### Interpretation

- VIF < 5 : Low multicollinearity
- 5 ≤ VIF < 10 : Moderate multicollinearity
- VIF ≥ 10 : High multicollinearity

# Heteroscedasticity

The Breusch–Pagan test is used to detect heteroscedasticity.

```{r}
hetero_check(fit)
```

### Interpretation

- p-value > 0.05 : Homoscedasticity assumption is satisfied.
- p-value ≤ 0.05 : Evidence of heteroscedasticity.

# Influential Observations

Cook's Distance is used to identify influential observations.

```{r}
outlier_check(fit)
```

### Interpretation

Observations with Cook's Distance greater than the threshold are considered influential and may have a strong impact on the fitted regression model.

# Overall Diagnostic Summary

The package also provides a concise summary of all diagnostic tests.

```{r}
model_diagnostics(fit)
```

This function summarizes:

- Normality
- Homoscedasticity
- Multicollinearity
- Influential observations

making it easier to evaluate the overall adequacy of the regression model.

# Conclusion

PrakKomstatPert8 provides an easy-to-use workflow for evaluating the assumptions of linear regression models.

The package includes functions for:

- Residual normality assessment
- Multicollinearity detection
- Heteroscedasticity testing
- Influential observation detection
- Overall diagnostic summary

These diagnostic tools help users assess whether a regression model satisfies the assumptions required for reliable statistical inference.
